---
title: "Introduction to Probability - Chapter 1"
output: html_notebook
---

## Vectors
How to create vectors
```{r}
# create vectors
v = c(3,1,4,1,5,9)
sum(v)
max(v)
length(v)
```
How to access a vectors
```{r}
v[2]
v[c(1,3)]
v[-(2:4)]
```

A lot of R operations are interpreted element-wise. For example $$(1, \frac{1}{2^2}, \dots, \frac{1}{100^2})$$ could be generated as follows.
```{r}
1/(1:100)^2
```

## Factorials and binomial coefficients

Factorial is computed using ```factorial(n)``` and $$n \choose k$$ is computed using ```choose(n,k)```. These functions also have logarithmic counterparts ```lfactorial(n)``` and ```lchoose(n,k)```

## Sampling and simulation

```sample``` command is used to draw a random sample. For example 
```{r}
n = 10
k = 5
sample(10,5) # without replacement
sample(10,5, replace = TRUE) # with replacement
```
Hence to generate a random permutation of $n$ numbers we can use ```sample(n,n)``` or, equivalently, ```sample(n)```
```{r}
sample(10,10)
sample(10)
```

```{r}
sample(letters, 7, replace = TRUE)
sample(4,3, replace = TRUE, prob = c(0.1,0.2,0.3,0.4))
```

## Problem Simulation

R provides functions that allow us simulate probability problems. In this section we will describe Matching problem simulation.

*de montmort's matching problem*: Consider a well-shuffled deck of n cards, labeled from 1 through n. While you are flipping cards one by one you and saying outloud the number from 1 to n, if the card's label matches with the number then you win the game. What is the probability of winning the game?

The probabilty of winning the game is $1 - \frac{1}{e}$. Let us show by simulation that this is true. 

```{r}
n = 100; # number of cards
r = replicate(10^4, sum(sample(n) == (1:n))); # shuffle and count the cards for 10,000 times
sum(r>=1)/10^4; # compute the empirical probability of winning
cat('1 - 1/e=', 1-1/exp(1));
```

