---
title: "Introduction to Probability - Chapter 2"
output: html_notebook
---

## Simulating the frequentist interpretation
According to the frequentist interpretation of probability, $P(A|B) = \frac{n_{AB}}{n_B}$, where $n_{AB}$ is the number of times A and B events happen together, $n_B$ is the number of times that event B happens.

*Elder is a girl vs. at least one girl*: A family has two children. What is the probability of having both girl, given the elder is a girl? What is the probability of having both girl, given at least one is a girl?

*Solution*: Let G: girl, B: boy then our sample space is $\{GG, GB,BG,BB\}$. 

P(both girl | elder is a girl) = P(both girls, elder is a girl)/ P(elder is girl) = (1/4)/(2/4) = 1/2.

P(both girl | at least one girl) = P(both girls, at least one girl) / P(at least one girl) = (1/4)/(3/4) = 1/3.

Let's confirm the above analysis using simulation. First let us prepare some data.
```{r}
n = 10^5; # number of families
child1 = sample(2, n, replace = TRUE); # first child gender 1: girl, 2: boy
child2 = sample(2, n, replace = TRUE) 
```

Let's compute P(both girl | elder is a girl)
```{r}
n.b = sum(child1 == 1)
n.ab = sum(child1 == 1 & child2 == 1)
cat(n.ab/n.b)
```
So it confirms our theorical analysis, i.e., the simulation result is very close to 1/2. What about P(both girl | at least one girl)

```{r}
n.b = sum(child1 == 1 | child2 == 1)
n.ab = sum(child1 == 1 & child2 == 1)
cat(n.ab/n.b)
```

Again the simulation result is close to our theoritical analysis.


## Monty Hall simulation (Conditioning as a problem-solving tool)
This is a game show, hosted by Monty. There are 3 doors and behind them there are 2 cows and 1 car.
The contesant choose 1 door. Monty then pick a door randomly with equal probability but he will never reveal the car. Monty then offers the contestant a chance to change his mind and pick another door. Should the contestant change the door if his/her aim is to get the car.

The user should adopt switch strategy since if he switches P(get car) = 2/3. The intuition is that if we imagine playing this game for 1000 times. Typically 333 times your intial guess will be correct. So if you adopt switch-strategy you will get the car the other 667 times. Let's simulate it

```{r}
n = 10^5 # number of playing times
cardoor = sample(3, n, replace = TRUE) # the door that the car is be behind
cat(sum(cardoor == 1)/n) 
```
Here we assume that the contestant always pick the door 1. This is reasonable since we can always rename the door. The result so that P(get a car if never switch) = 1/3
